---
title: "Lesson 02: Charts and Graphs"
output: html_notebook
---

References
• Black, Chapter 2 Charts and Graphs (pp. 15-34)
• Kabakoff, Chapter 2.5 Useful Functions (pp. 43-44)
• Chapter 6 Basic Graphs (pp. 117-136) 
• Davies, Chapter 14 Basic Data Visualization (pp. 298-308)
• Stowell, Chapter 8 Creating Plots (pp. 99-117)

Data set: home_prices.csv

Description: This data file is derived from a random sample of home resale records maintained by realtors.
There are 117 observations and eight variables:

1. PRICE = Selling price ($hundreds)
2. SQFT = Square feet of living space
3. YEAR = Year of construction (year)
4. BATHS = Number of bathrooms
5. FEATS = Number out of 11 features (dishwasher, refrigerator, microwave, disposal, washer, intercom, skylight(s), compactor, dryer, handicap fit, cable TV access)
6. NBR = Located in northeast sector of city (YES) or not (NO)
7. CORNER = Corner location (YES) or not (NO)
8. TAX = Annual taxes ($)

```{r}
#Read the comma-delimited text file creating a data frame object in R
#as we did in Lesson 1, then examine its structure:
houses <- read.csv("home_prices.csv") 
str(houses)
```

Exercises:
1) For the following exercises use hist(), plot(), boxplot() and par() functions supplied by R. a) Construct a histogram for PRICE. Describe the distribution shape.
```{r}
hist(houses$PRICE) # looks positively skewed, mean > median
```

```{r}
#or use with() function
#with(houses, hist(PRICE))
```


```{r}
hist(houses$TAX) # also looks positively skewed, mean > median
# with(houses, hist(TAX))
```
b) Construct a histogram for TAX. Describe the distribution shape.

```{r}

```

￼1
Histogram of houses$PRICE
￼￼￼￼￼￼￼￼￼￼￼1000 2000 3000 4000 5000
houses$PRICE
b) Construct a histogram for TAX. Describe the distribution shape.
￼￼￼￼￼￼￼￼￼￼￼1000 2000 3000 4000
houses$TAX
c) Construct a scatterplot displaying TAX versus PRICE. Is there a relationship?
￼# or use with() function

2
Frequency Frequency
0 10 20 30 0 10 20 30
￼plot(houses$PRICE, houses$TAX)
￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼2000 3000 4000 5000
houses$PRICE
d) Construct a stem-and-leaf plot for TAX using stem().
stem(houses$TAX)
￼# or use with() function
# with(houses, plot(PRICE, TAX))
￼##
##   The decimal point is 2 digit(s) to the right of the |
##
## 4|
## 6|
## 8|
## 10|
## 12|
## 14|
## 16|
## 18|
## 20|
## 22|
## 24|
## 26|
## 28|
## 30|
## 32|2
## 34|
## 36|2
## 38|345
## 40|69
## 42|93
## 44|1
66
648
03778022459
023688958
266889000006789
00222347788344557
011336888259
00333455767779
05581356
6339
9938
56078
286
3
houses$TAX
1000 3000
e) Use the par() and mfrow() or mfcol() functions to construct a multi-plot figure of two rows and one column showing the histograms for PRICE and TAX.
￼par(mfrow=c(1,2)) with(houses, hist(PRICE)) with(houses, hist(TAX))
Histogram of PRICE
Histogram of TAX
￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼￼1000 3000 5000
PRICE
1000 3000
TAX
￼par(mfrow=c(1,1))
2) For the following exercises use hist() and, within hist(), breaks.
a) Construct a histogram for PRICE starting the first class at 1300 with a class width of 600.
max(houses$PRICE) # this will let us know how many breaks we need ## [1] 5375
￼￼with(houses, hist(PRICE, breaks = c(1300, 1900, 2500, 3100, 3700, 4300, 4900, 5500)))
4
Frequency
0 10 20 30
Frequency
0 10 20 30
Histogram of PRICE
￼￼￼￼￼￼￼￼￼2000 3000 4000 5000
PRICE
b) Construct a histogram for TAX starting the first class at 500 with a class width of 500.
￼max(houses$TAX) ## [1] 4412.5
￼with(houses, hist(TAX, breaks = c(500, 1000, 1500, 2000, 2500, 3000, 3500, 4000, 4500)))
Histogram of TAX
￼￼￼￼￼￼￼￼￼￼1000 2000
3000 4000
TAX
5
Frequency Frequency
0 10 20 30 0 10 20 30 40
